---
title: "Integrative Analysis of GWAS Risk Genes in the Human Protein Atlas"
author: "Zhuoran Cai"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  html_document:
    self_contained: false
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("Functions.R")

#library(hrbrthemes)
library(viridis)
library(umap)
library(ggplot2)
library(dplyr)
library(MASS)
library(rmarkdown)
library(httr)
library(jsonlite)
library(tidyr)
library(plotly)
library(data.table)
library("ggVennDiagram")
library(VennDiagram)
library(ggvenn)
library(ggVennDiagram)
```

```{r,include=FALSE}
#import previous search result

.gwas_path    <- "data/search_result/gwas_list_FTD.RDS"
disease_data  <- readRDS(.gwas_path)
.disease_abbr <- sub(".*gwas_list_(.*)\\.RDS$", "\\1", .gwas_path)
.dc_map       <- c(AD = "#e84935", DLB = "#4fbdd7", FTD = "#00a087")
disease_color <- unname(.dc_map[.disease_abbr])
if (is.na(disease_color)) disease_color <- "steelblue"

HPA_data<-readRDS('data/search_result/HPA_data_FTD.RDS')

# Auto-detect HPA version from metadata stored during search
.hpa_url    <- attr(HPA_data, "HPA_version")
.hpa_raw    <- sub("https://(.*?)\\.proteinatlas.*", "\\1", .hpa_url)
HPA_VERSION <- if (.hpa_raw == "www") "latest" else .hpa_raw
hpa_dir     <- paste0("data/", HPA_VERSION, "/")

#import UMAP data and classification
UMAP_data<-read.delim(paste0(hpa_dir,'UMAP_and_cluster.',HPA_VERSION,'.txt'),sep = "\t", header=TRUE)
hpa_class_file    <- paste0(hpa_dir, 'HPA_classification.',    HPA_VERSION, '.csv')
hpa_class_sc_file <- paste0(hpa_dir, 'HPA_classification_SC.', HPA_VERSION, '.csv')
hpa_class_available <- file.exists(hpa_class_file) && file.exists(hpa_class_sc_file)

if (hpa_class_available) {
  HPA_classification    <- read.csv(hpa_class_file)
  HPA_classification_SC <- read.csv(hpa_class_sc_file)
} else {
  HPA_classification    <- NULL
  HPA_classification_SC <- NULL
}

#read dictionary
cluster_dict_brain<-read.csv(paste0(hpa_dir,'cluster_dict_brain.',HPA_VERSION,'.csv'))
cluster_dict_sc<-read.csv(paste0(hpa_dir,'cluster_dict_sc.',HPA_VERSION,'.csv'))
cluster_dict_tissue<-read.csv(paste0(hpa_dir,'cluster_dict_tissue.',HPA_VERSION,'.csv'))
```

```{r,include=FALSE}
identifier<- "ensembl_id"
new_data <- merge(UMAP_data, disease_data, by.x = "gene", by.y = identifier)
new_data<- merge(new_data,HPA_data, by.x = "gene", by.y = 'Ensembl')

new_data_filt<-new_data%>%group_by(gene) %>%
  slice(1) %>%
  ungroup()

if (hpa_class_available) {
  HPA_classification <- merge(HPA_classification, HPA_classification_SC, by = "ENSG")
}
```

This report is generated automatically by the **HPA GeneSet Explorer** pipeline. It maps GWAS-derived risk genes onto the Human Protein Atlas (HPA) to identify which cell types, tissues, and brain regions show significant enrichment of disease-associated gene expression.

The pipeline proceeds in three steps:

1.  Risk genes are retrieved from the **GWAS Catalog** using trait-specific EFO identifiers.
2.  Gene expression and cluster assignments are retrieved from the **Human Protein Atlas (HPA)**.
3.  Over-representation analysis and specificity profiling are performed to identify biologically relevant patterns.

------------------------------------------------------------------------

## 0. Search Parameters

```{r}
meta <- attr(disease_data, "search_efo_meta")
df_traits <- data.frame(
  "#"       = seq_len(nrow(meta)),
  "Trait"   = meta$trait,
  "EFO ID"  = meta$efo_id,
  check.names = FALSE
)
knitr::kable(df_traits, align = c("c", "l", "l"))
```

```{r, results='asis'}
gwas_date <- attr(disease_data, "search_date")
hpa_date  <- attr(HPA_data,     "search_date")
cat(paste0("**GWAS search date:** ",
  if (!is.null(gwas_date)) format(gwas_date, "%Y-%m-%d %H:%M:%S") else "N/A", "  \n"))
cat(paste0("**HPA version:** ", HPA_VERSION, "  \n"))
cat(paste0("**HPA search date:** ",
  if (!is.null(hpa_date)) format(hpa_date, "%Y-%m-%d %H:%M:%S") else "N/A", "  \n"))
cat(paste0("*(Source: ", attr(HPA_data, "HPA_version"), ")*\n"))
```

## 1. Summary of Statistics

This section gives a quick overview of how many risk genes were identified, how many were found in HPA, and how many could be mapped onto the brain expression UMAP.

```{r}
df_stats <- data.frame(
  "Description" = c(
    "Total genes in GWAS gene list",
    "Risk genes found in HPA",
    "Risk genes mappable to brain UMAP"
  ),
  "Count" = c(
    length(unique(disease_data$ensembl_id)),
    length(unique(HPA_data$Ensembl)),
    nrow(new_data_filt)
  ),
  check.names = FALSE
)
knitr::kable(df_stats, align = c("l", "r"))
```

## 2. Cluster Enrichment Analysis

HPA organises genes into expression clusters using UMAP-based dimensionality reduction. Each cluster groups genes with similar expression patterns across tissues, single-cell types, or brain regions. This section asks: **are disease risk genes over-represented in specific clusters?**

### 2a. Over-representation analysis (ORA) using Fisher's exact test

For each cluster, we test whether risk genes occur more often than expected by chance using Fisher's exact test. We then quantify enrichment using the **Fold Enrichment** score:

$$
\text{Fold Enrichment} = \frac{\text{Observed Overlap}}{\text{Expected Overlap}}, \quad \text{where} \quad \text{Expected Overlap} = \frac{\text{Risk Genes} \times \text{Cluster Size}}{\text{Total Genes}}
$$

-   **Fold Enrichment \> 1**: More risk genes in this cluster than expected - enriched.
-   **Fold Enrichment = 1**: No enrichment.
-   **Fold Enrichment \< 1**: Fewer risk genes than expected - depleted.

In the lollipop plots below, each dot represents one expression cluster. The x-axis shows fold enrichment and the y-axis shows the cluster label. **Orange dots** indicate clusters with a nominally significant enrichment (raw p \< 0.05). Genes not found in the HPA brain expression database are excluded from further analysis.

#### Raw p-value (p \< 0.05)

```{r include=FALSE}
total_genes <- length(UMAP_data$gene)
risk_genes <- length(new_data_filt$gene)

fisher_result_tissue<-(run_fisher_test(
  cluster_column ="UMAP.Cluster.tissue",
  p_threshold = 1,cluster_type = 'Tissue'
))

fisher_result_SC<-(run_fisher_test(
  cluster_column = "UMAP.Cluster.singlecell",
  p_threshold = 1,cluster_type = 'SingleCell'
))

fisher_result_brain<-(run_fisher_test(
  cluster_column = "UMAP.Cluster.brain",cluster_type = 'Brain',
  p_threshold = 1
))

fisher_result_tissue <- fisher_result_tissue %>%
  mutate(fold_enrichment = overlap / ((risk_genes * cluster_size) / total_genes))
fisher_result_SC <- fisher_result_SC %>%
  mutate(fold_enrichment = overlap / ((risk_genes * cluster_size) / total_genes))
fisher_result_brain <- fisher_result_brain %>%
  mutate(fold_enrichment = overlap / ((risk_genes * cluster_size) / total_genes))

fisher_result_tissue<- fisher_result_tissue %>%
  mutate(fisher_p_adj = p.adjust(fisher_p, method = "BH"))
fisher_result_SC<- fisher_result_SC%>%
  mutate(fisher_p_adj = p.adjust(fisher_p, method = "BH"))
fisher_result_brain <- fisher_result_brain %>%
  mutate(fisher_p_adj = p.adjust(fisher_p, method = "BH"))

disease_label <- attr(disease_data, "search_efo_meta")$trait[1]
```

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_tissue, fisher_result_tissue$fisher_p, 0.05,
              "Fold Enrichment Across Tissue Clusters (raw p < 0.05)")
```

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_SC, fisher_result_SC$fisher_p, 0.05,
              "Fold Enrichment Across Single Cell Clusters (raw p < 0.05)")
```

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_brain, fisher_result_brain$fisher_p, 0.05,
              "Fold Enrichment Across Brain Clusters (raw p < 0.05)")
```

#### After Benjamini-Hochberg (BH) multiple testing correction (adjusted p \< 0.1)

Because many clusters are tested simultaneously, we apply the Benjamini-Hochberg (BH) procedure to control the false discovery rate (FDR). The adjusted p-value threshold used here is 0.1, meaning we accept that up to 10% of flagged clusters may be false positives. Clusters passing this stricter threshold are considered more robustly enriched.

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_tissue, fisher_result_tissue$fisher_p_adj, 0.1,
              "Fold Enrichment Across Tissue Clusters (BH-adjusted p < 0.1)")
```

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_SC, fisher_result_SC$fisher_p_adj, 0.1,
              "Fold Enrichment Across Single Cell Clusters (BH-adjusted p < 0.1)")
```

```{r fig.width=9, fig.height=12}
plot_lollipop(fisher_result_brain, fisher_result_brain$fisher_p_adj, 0.1,
              "Fold Enrichment Across Brain Clusters (BH-adjusted p < 0.1)")
```

### 2b. UMAP visualisation of enriched clusters

Each point in the UMAP plots below represents one gene, positioned according to its global expression pattern. Genes with similar expression profiles cluster together in this two-dimensional space.

-   **Grey points**: all genes in the HPA database
-   **Coloured points**: genes belonging to clusters significantly enriched for risk genes (raw p \< 0.05)
-   **Cyan triangles**: risk genes located within enriched clusters

```{r,fig.width=10,fig.height=6}
UMAP_data_tissue<-read.csv(paste0(hpa_dir,'umap_tissue_data.',HPA_VERSION,'.csv'), header=TRUE)

significant_clusters <- fisher_result_tissue %>%
  filter(fisher_p < 0.05) %>%
  pull(UMAP.Cluster.tissue)

UMAP_data_tissue_with_enrichment <- UMAP_data_tissue %>%
  left_join(fisher_result_tissue %>% filter(fisher_p < 0.05) %>% select(UMAP.Cluster.tissue, fold_enrichment),
            by = c("cluster_no" = "UMAP.Cluster.tissue"))
ggplot() +
  geom_point(data = UMAP_data_tissue_with_enrichment,
             aes(x = x, y = y),
             size = 1.3, alpha = 0.4, shape = 16, color = "lightgrey") +
  geom_point(data = UMAP_data_tissue_with_enrichment %>%
               filter(cluster_no %in% significant_clusters),
             aes(x = x, y = y,
                 color = paste0("Cluster ", cluster_no, ": ", location_name)),
             size = 1.3, shape = 16) +
  geom_point(data = UMAP_data_tissue_with_enrichment %>%
               filter(ensembl_gene_id %in% new_data_filt$gene,
                      cluster_no %in% significant_clusters),
             aes(x = x, y = y),
             shape = 24, fill = "cyan3", size = 1.3, stroke = 0.7) +
  scale_color_discrete(name = "Enriched clusters") +
  theme_void() +
  labs(title = paste0("UMAP - Tissue expression (HPA ", HPA_VERSION, ")"))
```

```{r,fig.width=10,fig.height=6}
UMAP_data_SC <- read.csv(paste0(hpa_dir,'umap_singlecell_data.',HPA_VERSION,'.csv'), header = TRUE)

significant_clusters_SC <- fisher_result_SC %>%
  filter(fisher_p < 0.05) %>%
  pull(UMAP.Cluster.singlecell)

UMAP_data_SC_with_enrichment <- UMAP_data_SC %>%
  left_join(
    fisher_result_SC %>%
      filter(fisher_p < 0.05) %>%
      select(UMAP.Cluster.singlecell, fold_enrichment),
    by = c("cluster_no" = "UMAP.Cluster.singlecell")
  )

ggplot() +
  geom_point(data = UMAP_data_SC_with_enrichment,
             aes(x = x, y = y),
             size = 1.3, alpha = 0.4, shape = 16, color = "lightgrey") +
  geom_point(data = UMAP_data_SC_with_enrichment %>%
               filter(cluster_no %in% significant_clusters_SC),
             aes(x = x, y = y,
                 color = paste0("Cluster ", cluster_no, ": ", location_name)),
             size = 1.3, shape = 16) +
  geom_point(data = UMAP_data_SC_with_enrichment %>%
               filter(ensembl_gene_id %in% new_data_filt$gene,
                      cluster_no %in% significant_clusters_SC),
             aes(x = x, y = y),
             shape = 24, fill = "cyan3", size = 1.3, stroke = 0.7) +
  scale_color_discrete(name = "Enriched clusters") +
  theme_void() +
  labs(title = paste0("UMAP - Single-cell expression (HPA ", HPA_VERSION, ")"))
```

```{r,fig.width=10,fig.height=6}
UMAP_data_brain <- read.csv(paste0(hpa_dir,'umap_brain_data.',HPA_VERSION,'.csv'), header = TRUE)

significant_clusters_brain <- fisher_result_brain %>%
  filter(fisher_p < 0.05) %>%
  pull(UMAP.Cluster.brain)

UMAP_data_brain_with_enrichment <- UMAP_data_brain %>%
  left_join(
    fisher_result_brain %>%
      filter(fisher_p < 0.05) %>%
      select(UMAP.Cluster.brain, fold_enrichment),
    by = c("cluster_no" = "UMAP.Cluster.brain")
  )

ggplot() +
  geom_point(data = UMAP_data_brain_with_enrichment,
             aes(x = x, y = y),
             size = 1.3, alpha = 0.4, shape = 16, color = "lightgrey") +
  geom_point(data = UMAP_data_brain_with_enrichment %>%
               filter(cluster_no %in% significant_clusters_brain),
             aes(x = x, y = y,
                 color = paste0("Cluster ", cluster_no, ": ", location_name)),
             size = 1.3, shape = 16) +
  geom_point(data = UMAP_data_brain_with_enrichment %>%
               filter(ensembl_gene_id %in% new_data_filt$gene,
                      cluster_no %in% significant_clusters_brain),
             aes(x = x, y = y),
             shape = 24, fill = "cyan3", size = 1.3, stroke = 0.7) +
  scale_color_discrete(name = "Enriched clusters") +
  theme_void() +
  labs(title = paste0("UMAP - Brain expression (HPA ", HPA_VERSION, ")"))
```

### 2c. Monte Carlo enrichment analysis

As a complement to Fisher's exact test, we validate cluster enrichment using a Monte Carlo simulation. For each UMAP cluster type, risk genes are randomly redistributed across clusters 1,000,000 times (can be adjusted), proportional to cluster size. The Z-score measures how many standard deviations the observed overlap deviates from this null distribution. Highlighted clusters (coloured stems and points) have empirical p \< 0.05 and a positive Z-score (enriched, not depleted).

**Parameter setting**

```{r,include=TRUE,echo=TRUE}
n_iter <- 10000  # paper uses 1,000,000; reduced here for faster rendering
```

```{r include=FALSE}
mc_result_tissue <- run_mc(fisher_result_tissue, "UMAP.Cluster.tissue",      n_iter)
mc_result_SC     <- run_mc(fisher_result_SC,     "UMAP.Cluster.singlecell",  n_iter)
mc_result_brain  <- run_mc(fisher_result_brain,  "UMAP.Cluster.brain",       n_iter)
```

```{r fig.width=5, fig.height=max(7, ceiling(nrow(mc_result_tissue) * 0.1))}
plot_mc_lollipop(mc_result_tissue, "UMAP.Cluster.tissue",
                 disease_color, paste0("Tissue - ", disease_label))
```

```{r fig.width=5, fig.height=max(7, ceiling(nrow(mc_result_SC) * 0.1))}
plot_mc_lollipop(mc_result_SC, "UMAP.Cluster.singlecell",
                 disease_color, paste0("Single-cell - ", disease_label))
```

```{r fig.width=5, fig.height=max(7, ceiling(nrow(mc_result_brain) * 0.1))}
plot_mc_lollipop(mc_result_brain, "UMAP.Cluster.brain",
                 disease_color, paste0("Brain - ", disease_label))
```

### 2d. Volcano Plots

The volcano plots below provide a complementary view of the enrichment results. Each point is one cluster. The x-axis shows fold enrichment and the y-axis shows statistical significance (-log of BH-adjusted p-value). Points further right are more enriched; points higher up are more significant.

-   **Grey**: not significant at raw p \< 0.05
-   **Orange**: significant at raw p \< 0.05
-   **Red**: significant after BH correction (adjusted p \< 0.1)
-   **Dashed line**: fold enrichment = 1 (no enrichment)

The table below each plot lists the clusters that passed BH-adjusted p \< 0.1.

```{r}
ggplot(fisher_result_tissue, aes(fold_enrichment, -log(fisher_p_adj))) +
  geom_point(color = "grey70", alpha = 0.6) +
  geom_point(data = subset(fisher_result_tissue, fisher_p < 0.05), color = "orange", size = 2) +
  geom_point(data = subset(fisher_result_tissue, fisher_p_adj < 0.1), color = "red", size = 2) +
  geom_text(data = subset(fisher_result_tissue, fisher_p < 0.05),
            aes(label = UMAP.Cluster.tissue), vjust = -0.5, size = 3) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "skyblue3") +
  labs(title = paste0("Tissue clusters - ", disease_label),
       x = "Fold Enrichment", y = "-log(BH-adjusted p-value)")

print(fisher_result_tissue %>% filter(fisher_p_adj < 0.1))

ggplot(fisher_result_SC, aes(fold_enrichment, -log(fisher_p_adj))) +
  geom_point(color = "grey70", alpha = 0.6) +
  geom_point(data = subset(fisher_result_SC, fisher_p < 0.05), color = "orange", size = 2) +
  geom_point(data = subset(fisher_result_SC, fisher_p_adj < 0.1), color = "red", size = 2) +
  geom_text(data = subset(fisher_result_SC, fisher_p < 0.05),
            aes(label = UMAP.Cluster.singlecell), vjust = -0.5, size = 3) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "skyblue3") +
  labs(title = paste0("Single-cell clusters - ", disease_label),
       x = "Fold Enrichment", y = "-log(BH-adjusted p-value)")

print(fisher_result_SC %>% filter(fisher_p_adj < 0.1))

ggplot(fisher_result_brain, aes(fold_enrichment, -log(fisher_p_adj))) +
  geom_point(color = "grey70", alpha = 0.6) +
  geom_point(data = subset(fisher_result_brain, fisher_p < 0.05), color = "orange", size = 2) +
  geom_point(data = subset(fisher_result_brain, fisher_p_adj < 0.1), color = "red", size = 2) +
  geom_text(data = subset(fisher_result_brain, fisher_p < 0.05),
            aes(label = UMAP.Cluster.brain), vjust = -0.5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "skyblue3") +
  labs(title = paste0("Brain clusters - ", disease_label),
       x = "Fold Enrichment", y = "-log(BH-adjusted p-value)")

print(fisher_result_brain %>% filter(fisher_p_adj < 0.1))
```

## 3. Expression Specificity of Risk Genes

Beyond cluster enrichment, we examine how selectively risk genes are expressed. The Human Protein Atlas classifies each gene into specificity categories based on how its expression compares across tissues or cell types.

**Elevated expression categories** (source: <https://www.proteinatlas.org/humanproteome/tissue/tissue+specific>):

| Category | Definition |
|------------------------------------|------------------------------------|
| **Tissue enriched** | At least 4-fold higher mRNA level in one tissue compared to all others |
| **Group enriched** | At least 4-fold higher average mRNA level in a group of 2-5 tissues compared to all others |
| **Tissue enhanced** | At least 4-fold higher mRNA level in one tissue compared to the average of all other tissues |
| **Low specificity** | Does not meet any of the above thresholds - broadly expressed |

We examine how risk genes are distributed across these categories in tissues, single-cell types, and brain regions.

```{r,results='asis'}
if (!hpa_class_available) {
  cat(paste0(
    "\n> **Data unavailable for HPA ", HPA_VERSION, ":** ",
    "`HPA_classification.", HPA_VERSION, ".csv` and ",
    "`HPA_classification_SC.", HPA_VERSION, ".csv` ",
    "have not yet been added to `data/", HPA_VERSION, "/`. ",
    "Sections 3a and 3b are skipped. ",
    "Please add the files and re-run the report.\n"
  ))
}
```

### 3a. Specificity distribution of risk genes

The pie charts below show how risk genes are distributed across specificity categories, across three expression contexts. Each slice corresponds to one specificity class; the percentage indicates the proportion of risk genes in that class.

**Top row** (from HPA search result): tissue RNA specificity, single-cell type specificity, single-nuclei brain specificity. **Bottom row** (from HPA classification file): brain regional, brain domain, brain subregional specificity.

```{r,fig.width=10,fig.height=9,eval=hpa_class_available}
risk_list <- new_data_filt$gene
risk_classification <- HPA_classification[HPA_classification$ENSG %in% risk_list, ]

color_map <- function(labels){
  lbl <- tolower(labels)
  ifelse(lbl == "group enriched", "violetred3",
  ifelse(grepl("enriched", lbl), "violetred4",
  ifelse(grepl("enhanced", lbl), "pink2",
  ifelse(grepl("low", lbl), "mistyrose2", "grey"))))
}

plot_spec <- function(df, cols){
  for (col in cols) {
    counts <- table(df[[col]])
    perc   <- round(100 * counts / sum(counts), 1)
    labels <- paste0(names(counts), " (", perc, "%)")
    pie(counts, main = col, col = color_map(names(counts)), labels = labels)
  }
}

par(mfrow = c(3, 2))
plot_spec(new_data_filt, c("RNA.tissue.specificity","RNA.single.cell.type.specificity","RNA.single.nuclei.brain.specificity"))
plot_spec(risk_classification, c("Human.brain.regional.RNA.Specificity","Human.brain.domain.RNA.Specificity","Human.brain.subregional.RNA.Specificity"))
```

### 3b. Tissue distribution of tissue-enriched risk genes

For risk genes classified as "Tissue enriched", this section shows which specific tissues they are enriched in. Each panel corresponds to one expression context. The x-axis shows the tissue name and the y-axis shows the count of risk genes enriched in that tissue. The red label indicates the total number of tissue-enriched genes in each context.

```{r fig.height=7,eval=hpa_class_available}
combined_data <- bind_rows(
  risk_classification %>%
    filter(Human.brain.regional.RNA.Specificity == "Tissue enriched") %>%
    mutate(Category = "Human Brain Regional enriched", Enrichment_Tissue = Human.brain.regional.Enriched.Enhanced.Tissue),
  risk_classification %>%
    filter(Consensus.RNA.Specificity == "Tissue enriched") %>%
    mutate(Category = "Consensus RNA Tissue enriched", Enrichment_Tissue = Consensus.Enriched.Enhanced.Tissue),
  risk_classification %>%
    filter(Human.brain.domain.RNA.Specificity == "Tissue enriched") %>%
    mutate(Category = "Human Brain Domain enriched", Enrichment_Tissue = Human.brain.domain.Enriched.Enhanced.Tissue),
  risk_classification %>%
    filter(Human.brain.subregional.RNA.Specificity == "Tissue enriched") %>%
    mutate(Category = "Human Brain Subregional enriched", Enrichment_Tissue = Human.brain.subregional.Enriched.Enhanced.Tissue),
  risk_classification %>%
    filter(Single.cell.type.RNA.Specificity == "Tissue enriched") %>%
    mutate(Category = "Single Cell Type enriched", Enrichment_Tissue = Single.cell.type.Enriched.Enhanced.Tissue)
)

total_counts <- combined_data %>%
  group_by(Category) %>%
  summarise(Total = dplyr::n())

ggplot(combined_data, aes(x = Enrichment_Tissue)) +
  geom_bar(fill = "steelblue") +
  facet_wrap(~Category, scales = "free", ncol = 2) +
  labs(y = "Number of risk genes") +
  geom_text(data = total_counts,
            aes(x = -Inf, y = Inf, label = paste("Total:", Total)),
            hjust = -0.1, vjust = 1.5, inherit.aes = FALSE, size = 3, color = "red") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 3c. Overlap between enriched gene sets (Venn diagrams)

The Venn diagrams below show how many risk genes are classified as enriched (tissue enriched or group enriched) across multiple expression contexts simultaneously. Overlapping regions indicate genes enriched in more than one context.

**Diagram 1**: overlap across tissue RNA, single-cell type, and single-nuclei brain contexts (from HPA search result).

```{r,fig.width=7,fig.height=5}
library(VennDiagram)

venn_list <- list(
  "Tissue Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.tissue.specificity %in% c("Tissue enriched", "Group enriched")]),
  "Single Cell Type Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.single.cell.type.specificity %in% c("Cell type enriched", "Group enriched")]),
  "Single Nuclei Brain Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.single.nuclei.brain.specificity %in% c("Cell type enriched", "Group enriched")])
)

ggvenn(
  venn_list,
  fill_color = c("#0073C2FF", "#EFC000FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 4, text_size = 3
) + ggtitle("Overlap of enriched risk genes across expression contexts")
```

**Diagram 2**: overlap across tissue RNA, single-cell type, and brain regional contexts.

```{r,fig.width=7,fig.height=5}
venn_list2 <- list(
  "Tissue Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.tissue.specificity %in% c("Tissue enriched", "Group enriched")]),
  "Single Cell Type Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.single.cell.type.specificity %in% c("Cell type enriched", "Group enriched")]),
  "Brain Regional Enriched" = unique(new_data_filt$gene[new_data_filt$RNA.brain.regional.specificity %in% c("Region enriched", "Group enriched")])
)

ggvenn(
  venn_list2,
  fill_color = c("#0073C2FF", "#EFC000FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 4, text_size = 3
) + ggtitle("Overlap of enriched risk genes (tissue, single-cell, brain regional)")
```
